Bayesian Sparsification of $\cplx$-valued networks

### 2019-12-09 paper/paper.tex
use term 'arithmetic complexity'; fixed the use of intractable

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embedded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and numerical computational
complexity restrictions \citep{he_amc:_2018}. Therefore compression methods for achieving
high model sparsity and numerical efficiency without losing much in performance are
especially relevant.
% Solving the storage and compute constraints encourages development of model compression
% and sparsification methods, that focus on favourable tradeoff between performance and size.

Signal processing, specifically, audio or radio frequency signal analysis is inherently
complex-valued, yet $\cplx$-valued models and neural networks, and, in particular, compression
methods tailored to $\cplx$-domain remain a niche field of research. Computational efficiency
reduction methods such as quantization, integer-based arithmetic \citep{citation_needed},
and distillation \citep{distillation} appear to be immediately applicable to $\cplx$-field.
%
Some model sparsification methods, e.g. magnitude pruning \citep{zuo_compression_2019}, which
reactivates zeroed parameters based on gradients, and sparsity inducing regularizers, can be
adapted directly to $\cplx$-valued models. Other methods require additional considerations.
% the more explicitly zero parameters there is, the less computations is needed.

%%% >>>> (cplx)
% complex nets and related stuff
% cplx domain models
% (cplx-transformer) yang-ma-li.tex
% (cplx capacity) monning-manandhar.tex
\citet{monning_evaluation_2018} state,
that $\cplx$-networks have less intuitive architectures, inherently higher numerical
complexity due to multiplication in $\cplx$, that requires 3-4 floating point multiplications,
dependening on the implementation.
% due to them not being true cplx and just paired real computational graphs

@@

The posterior itself is computationally intractable, save for the relatively simple cases
\citep{citation_needed}, and therefore exact inference is traded for tractability and scalability
of Variational Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach
recasts posterior inference into variational optimization problem: finding $q$ that is close
to $p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \citep{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

@@

% criticism and discussion
based on the algebraic equivalence between $2 \times 2$ real matrices and complex numbers.
Due to this constraint $\cplx$-networks have less expressive power for the same computation
complexity compared to a $\real$-layer with twice as many inputs and outputs.
%
The \verify{most fierce critique} of Bayesian approach towards model
sparsification was voiced in \citep{gale_state_2019}, where it was show that it yielded
inconsistent results in text translation tasks and systematically underperformed simpler
techniques such as norm pruning and $\ell_0$-regularization, \citep{louizos_learning_2017}.

>>>>

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embedded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and arithmetic (computational)
complexity restrictions \citep{he_amc:_2018}. Therefore compression methods for achieving
high model sparsity and numerical efficiency without losing much in performance are
especially relevant.
% Solving the storage and artihmetic constraints encourages development of model compression
% and sparsification methods, that focus on favourable tradeoff between performance and size.

Signal processing, specifically, audio or radio frequency signal analysis is inherently
complex-valued, yet $\cplx$-valued models and neural networks, and, in particular, compression
methods tailored to $\cplx$-domain remain a niche field of research. Computational efficiency
improvement methods such as quantization, integer-based arithmetic \citep{citation_needed},
and distillation \citep{distillation} appear to be immediately applicable to $\cplx$-field.
%
Some model sparsification methods, e.g. magnitude pruning \citep{zuo_compression_2019}, which
reactivates zeroed parameters based on gradients, and sparsity inducing regularizers, can be
adapted directly to $\cplx$-valued models. Other methods require additional considerations.
% the more explicitly zero parameters there is, the less computations is needed.

%%% >>>> (cplx)
% complex nets and related stuff
% cplx domain models
% (cplx-transformer) yang-ma-li.tex
% (cplx capacity) monning-manandhar.tex
\citet{monning_evaluation_2018} state,
that $\cplx$-networks have less intuitive architectures, inherently higher arithmetic
complexity due to multiplication in $\cplx$, that requires 3-4 floating point multiplications,
dependening on the implementation.
% due to them not being true cplx and just paired real computational graphs

@@

The posterior itself is intractable, save for the relatively simple cases \citep{citation_needed},
and therefore exact inference is traded for tractability and scalability of Variational
Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach recasts
posterior inference into variational optimization problem: finding $q$ that is close to
$p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \citep{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \citep{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

@@

% criticism and discussion
based on the algebraic equivalence between $2 \times 2$ real matrices and complex numbers.
Due to this constraint $\cplx$-networks have less expressive power for the same arithmetic
complexity compared to a $\real$-layer with twice as many inputs and outputs.
%
The \verify{most fierce critique} of Bayesian approach towards model
sparsification was voiced in \citep{gale_state_2019}, where it was show that it yielded
inconsistent results in text translation tasks and systematically underperformed simpler
techniques such as norm pruning and $\ell_0$-regularization, \citep{louizos_learning_2017}.
